---
title: "Practice Lab 10"
author: "Sean Huvaya"
date: "2024-03-26"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
```
###	Loading the dataset 
```{r}
df <- read.csv("/Users/ktxdev/Documents/MS in AI/Computational Statistics and Probability/M10/Practice Lab 10.csv")
```

### Stratifying the sample population in tertiles of SBP
```{r}
df$SBP_tertiles <- ntile(df$SBP, 3)
df$SBP_tertiles_factor <- factor(df$SBP_tertiles, levels = c(1:3), labels = c("Tertile1","Tertile2","Tertile3"))
```

### Comparing the age of subjects in the respective tertiles utilizing ANOVA
```{r}
anova_test <- aov(df$age ~ df$SBP_tertiles_factor)
summary(anova_test)
```

### Stratifying the population in subjects younger that 35, between 35 and 65 and those older than 65 years
```{r}
df$age_cat <- NA
df$age_cat[df$age < 35] <- 1
df$age_cat[df$age >= 35 & df$age <= 65] <- 2
df$age_cat[df$age > 65] <- 3
df$age_cat_factor <- factor(df$age_cat, levels = c(1:3), labels = c("below35", "between35_65", "above65"))
```

###	Comparing the SBP (as a continuous variable) of subjects in the age categories utilizing ANOVA
```{r}
anova_test<-aov(df$SBP ~ df$age_cat_factor)
summary(anova_test)
```

###	Building a contingency table displaying the counts of SBP tertiles and age categories
```{r}
SBP_age_table <- table(df$age_cat_factor, df$SBP_tertiles_factor)
SBP_age_table
```
###	Testing for associations between these categories using the chi square test
```{r}
result <- chisq.test(SBP_age_table)
result
```
#### Observed Values
```{r}
result$observed
```
#### Expected Values
```{r}
result$expected
```
#### Residuals Values
```{r}
result$residuals
```
###	Calculate relative risk to be in the highest tertile of SBP for the oldest in the dataset compared to those younger than 65 years
```{r}
risk_young <- sum(SBP_age_table[1:2, 3]) / sum(SBP_age_table[1:2, 1:3])
risk_old <- sum(SBP_age_table[3, 3]) / sum(SBP_age_table[3, 1:3])
relative_Risk <- risk_old / risk_young
relative_Risk
```
