---
title: "Practice Lab 9"
author: "Sean Huvaya"
date: "2024-03-19"
output: html_document
---

```{r setup, include=FALSE}
library(pwr)
```

### Planned RCT sample sizes needed (sig level of 0.05; type =two- sided) to detect a difference with a power of 90% under the assumption of:
#### a)	a small effect size (0.2)
```{r}
pwr.t.test(power = 0.9, d = 0.2, sig.level = .05)
```

#### b)	a medium effect size (0.5)
```{r}
pwr.t.test(power = 0.9, d = 0.5, sig.level = .05)
```
#### c)	a large effect size (0.8).
```{r}
pwr.t.test(power = 0.9, d = 0.8, sig.level = .05)
```
#### d) The power at large effect size when 36 subjects are enrolled
```{r}
pwr.t.test(n = 36, d = 0.8, sig.level = .05)
```
### The sample sizes needed within each group when comparing 5 groups using a one-way ANOVA in a planned RCT (sig level of 0.05) aiming for a power of 90% with:
#### a) a small effect size (0.2)
```{r}
pwr.anova.test(k = 5, sig.level = .05, f = 0.2, power = 0.9)
```
#### b) a medium effect size (0.5)
```{r}
pwr.anova.test(k = 5, sig.level = .05, f = 0.5, power = 0.9)
```
#### c) a large effect size (0.8)
```{r}
pwr.anova.test(k = 5, sig.level = .05, f = 0.8, power = 0.9)
```
#### d) Tthe power when 40 subjects are enrolled?
```{r}
pwr.anova.test(n = 8, f = 0.8, sig.level = 0.05, k = 5)
```